---
title: "Graphing Options - GISAID and Data Portal Uploads"
author: "Erin Gill"
date: "28/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(reshape)
library(plyr)
library(dplyr)
library(viridis)
```

## Considerations

In order to incorporate information on each province's uploads to **GISIAD** and the **VirusSeq Data Portal**, we need to make a few decisions about: 

- *how to present the data*
- *whether we want to include historical data or just the most recent data* (I have historical data going back to June 2021 for the Data Portal and September 2020 for GISAID)

In order to make these decisions, I've made some sample plots with a test dataset. 

```{r load_data, include = FALSE}
setwd("~/Projects/gisaid_uploads")
seqs <- read.csv('data/days_to_deposition_90.csv', header = TRUE)

# relevel factors
seqs$country <- factor(seqs$country)
seqs$country <- relevel(seqs$country, "Canada")

# rename columns
names(seqs)[2] <- 'GISAID'
names(seqs)[5] <- 'Portal'
names(seqs)[6] <- 'Mar_2022'

seqs <- subset(seqs, select = -c(percent_cases_sequenced_and_shared, cases, Mar_2022))

```

### Let's take a look at the dataset 

```{r, echo = FALSE}
knitr::kable(seqs, caption = "Number of sequences uploaded to GISAID and the Data Portal by country")
```

### First, let's look at a bar plot

Sequences uploaded to the two repositories are overlaid for each country, and the alpha parameter is decreased so that you can easily see when a country has uploaded more to one repository than another

In the following plots, imagine that countries are **provinces**
```{r, echo=FALSE}
#plot of median days to deposition to GISAID over last 90 days
ggplot(seqs, aes(x = country, y = GISAID, fill = 'GISAID', alpha = 0.5)) +
  geom_col() +
  scale_alpha(guide='none') +
  geom_col(aes(x = country, y = Portal, fill = 'Portal', alpha = 0.5)) +
  ylab("Number of sequences uploaded") +
  ggtitle("Total number of sequences shared to GISAID and the Data\nPortal by country") +
  guides(x = guide_axis(angle = 90)) +
  geom_text(size=2.5, aes(y = GISAID, label = GISAID), vjust = 1.5, colour = "black") +
  geom_text(size=2.5, aes(y = Portal, label = Portal), vjust = 1.5, colour = "black") +
  labs(fill='Repository') +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  scale_fill_manual(values = c("red", "blue"))

```

### Here's a second barplot option where sequences uploaded to each repository are plotted side by side for each country

```{r, melt_df, echo = FALSE, include=FALSE}
seqsmelt <- melt(seqs)

names(seqsmelt)[2] <- 'type'
names(seqsmelt)[3] <- 'sequences'

```


```{r, echo=FALSE}
ggplot(seqsmelt, aes(x = country, y = sequences, fill = type)) +
  geom_bar(position ='dodge', stat = 'identity') +
  ylab("Number of sequences uploaded") +
  xlab("Country") +
  ggtitle("Total number of sequences shared to GISAID and the Data\nPortal by country") +
  guides(x = guide_axis(angle = 90)) +
  labs(fill='Repository') +
  scale_fill_viridis(discrete = TRUE) 

```

```{r, import_date_data, echo=FALSE}

seqs2 <- read.csv('data/sample_uploads.csv', header = TRUE)

# relevel factors
seqs2$country <- factor(seqs$country)
seqs2$country <- relevel(seqs$country, "Canada")
seqs2$date <- as.Date(seqs2$date)

```
## What if we want to look at upload progress over time?

Here's the first few rows of a second dataset that contains multiple time points
```{r, echo = FALSE}
knitr::kable(seqs2[1:10, ], caption = "Number of sequences uploaded to GISAID and the Data Portal over time by country")
```

### Here each country is plotted separately and you can see if the "rate of upload" has been increasing or decreasing to both repositories
```{r, echo = FALSE}
ggplot(seqs2, aes(x = date, y = number)) +
  geom_point(aes(color = repository)) +
  geom_line(aes(color = repository)) +
  facet_wrap(~country) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_date(breaks = as.Date(c("2021-08-01", "2021-11-01")),
               labels = c('Aug', 'Nov')) +
  guides(x = guide_axis(angle = 90)) +
  ylab("Number of sequences uploaded") +
  xlab("Country") +
  ggtitle("Number of sequences shared to GISAID and the Data\nPortal over time by country")

```

### The same plot as above, but a bar plot instead of a line plot
```{r, echo=FALSE}
ggplot(seqs2, aes(x = date, y = number, fill = repository)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  facet_wrap(~country, ncol = 4) +
  scale_fill_viridis(discrete = TRUE) +
  guides(x = guide_axis(angle = 90)) +
  scale_x_date(breaks = as.Date(c("2021-08-01", "2021-11-01")),
               labels = c('Aug', 'Nov')) +
  ylab("Number of sequences uploaded") +
  xlab("Country") +
  ggtitle("Number of sequences shared to GISAID and the Data\nPortal over time by country")

```


